---
title: "cegwas results"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    self_contained: no
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, cache = F}

library(knitr)
library(cegwas)
library(dplyr)
library(ggplot2)
library(jsonlite)
library(RMySQL)
library(tidyr)
library(readr)
library(compiler)


# Get Payload
if (length(commandArgs(trailingOnly=TRUE)) == 0) {
  args <- fromJSON('{"submission_date": "2016-02-28T01:11:21.888460-06:00", "trait_name": "test5", "success": true, "trait_slug": "test5", "report_hash": "6a930ca9daedfa86d2eb", "report_name": "test5", "version": 1, "release": 0, "report_slug": "test5", "email": "danielecook@gmail.com"}')
} else {
  args <- fromJSON(commandArgs(trailingOnly=TRUE))
}

mysql_credentials <- fromJSON(readLines("credentials.json"))

# To connect to a database first create a src:
db <- src_mysql(dbname = "cegwas_v2", host = mysql_credentials$host, user = mysql_credentials$user, password= mysql_credentials$password)


update_status <- function(status) {

db <- src_mysql(dbname = "cegwas_v2", host = mysql_credentials$host, user = mysql_credentials$user, password= mysql_credentials$password)

# Function for updating status of currently running job.
comm <- sprintf("UPDATE report_trait SET status='%s' WHERE report_slug = '%s' AND trait_slug = '%s'",
                status, args$report_slug, args$trait_slug)
dbSendQuery(db$con, comm)
try(invisible(dbDisconnect(db$con)), silent = T)
}

report_trait_strain_tbl <- tbl(db, "report_trait_strain_value")

# Then reference a tbl within that src
input_trait <- collect(report_trait_strain_tbl %>%
             dplyr::select(strain, report_slug, trait_slug, value) %>%
             dplyr::filter(trait_slug == args$trait_slug, report_slug == args$report_slug) %>%
             dplyr::select(strain, value))

try(invisible(dbDisconnect(db$con)), silent = T)

colnames(input_trait) <- c("strain", args$trait_slug)

output <- opts_knit$get("rmarkdown.pandoc.to")
opts_chunk$set(warning=F,
               message=F,
               echo=F,
               eval=T,
               cache=F,
               fig.path="figures/",
               cache.path="cache/",
               results="hide")




pub_theme <- ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(size=14, color="black"),
                 axis.text.y = ggplot2::element_text(size=14, color="black"),
                 axis.title.x = ggplot2::element_text(size=14, face="bold", color="black", vjust=-.3),
                 axis.title.y = ggplot2::element_text(size=14, face="bold", color="black", vjust=2),
                 strip.text.x = ggplot2::element_text(size=16, face="bold", color="black"),
                 strip.text.y = ggplot2::element_text(size=16, face="bold", color="black"),
                 axis.ticks= element_line( color = "black", size = 0.25), 
                 legend.position="none",
                 plot.margin = unit(c(1.0,0.5,0.5,1.0),"cm"),
                 plot.title = ggplot2::element_text(size=24, face="bold", vjust = 1),
                 legend.position="none",
                 panel.background = ggplot2::element_rect(color = "black", size= 0.75),
                 strip.background = ggplot2::element_rect(color = "black", size = 0.75)) 


opts_chunk$set(fig.width=12, fig.height=6)
output <- "html"
```

```{r perform_mapping}

update_status("Processing Phenotype Data")
trait  <- process_pheno(input_trait)

update_status("Performing Mapping")

gwas <- cmpfun(gwas_mappings)
system.time(mapping <- gwas(trait, mapping_snp_set = FALSE))

update_status("Processing Mapping")
pr_mapping <- dplyr::filter(mapping, log10p !=0) %>% filter(!(grepl("MtDNA",marker) & log10p < 5))

max_sig <- max(pr_mapping$log10p) 
  
bf <- -log10(.05/nrow(pr_mapping))


proc_mappings <- data.frame()
if(max_sig > bf){
  warning("max_sig")
  proc_mappings <- process_mappings(mapping,trait) %>% 
    dplyr::filter(log10p !=0) %>% 
    dplyr::mutate(marker = gsub("_", ":", marker)) %>%
    filter(!(grepl("MtDNA",marker) & log10p < BF))
  readr::write_tsv(proc_mappings, "tables/processed_sig_mapping.tsv")
} 

processed_phenotypes <- as.data.frame(t(trait[[2]]))

mapping %>% dplyr::mutate(marker = gsub("_",":", marker)) %>%
readr::write_tsv("tables/raw_mapping.tsv")

input_trait %>% readr::write_tsv("tables/raw_phenotype.tsv")

as.data.frame(t(trait[[2]])) %>% dplyr::add_rownames("strain") %>% plyr::rename(c("V1"=args$trait_name)) %>%
readr::write_tsv("tables/processed_phenotype.tsv")

```

# Phenotype {.tabset  .tabset-fade}

## Distribution 

```{r phenotype_histogram}
update_status("Plotting Figures")
phenotype_data <- tidyr::gather(data.frame(trait[[2]]), strain, value)
ggplot2::ggplot(phenotype_data, aes(x = value)) +
  ggplot2::geom_histogram(color = "#0A3872", fill = "#0B5DA2") +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(size=18, face="bold", color="black"),
                 axis.text.y = ggplot2::element_text(size=18, face="bold", color="black"),
                 axis.title.x = ggplot2::element_text(size=24, face="bold", color="black", vjust=-.3),
                 axis.title.y = ggplot2::element_text(size=24, vjust = -0.3,  color="black"),
                 strip.text.x = ggplot2::element_text(size=24, face="bold", color="black"),
                 strip.text.y = ggplot2::element_text(size=16, face="bold", color="black"),
                 plot.title = ggplot2::element_text(size=24, face="bold", vjust = 1, margin = margin(b = 20, unit = "pt")),
                 legend.position="none") +
  ggplot2::labs(x = args$trait_name, y = "Count")

phenotype_combined <- tidyr::gather(data.frame(trait[[2]]), strain, value) %>%
                      dplyr::left_join(input_trait)

```

`r if (nrow(proc_mappings) >= 0) '__Histogram representation of your phenotype__'`

`r if (nrow(proc_mappings) > 0) 'Your phenotype is significantly associated with genetic variation present in the _C. elegans_ population!'`

`r if (nrow(proc_mappings) == 0) 'Unfortunately your phenotype was not significantly associated with genetic variation present in the _C. elegans_ population. This could be due to noisy trait data - have you performed heritability analysis for your phenotype using our heritability strain panel? If you have and the heritability was found to be high for this trait, the trait might not have reached statistical significance because it is highly complex and more strains need to be phenotyped. If you have phenotyped the entire 152 wild-isolate collection, you can patiently wait for more isolates to be added to the collection, or generate an F2 recombinant inbred line (RIL) panel generated between strains with high and low phenotypes. '`

## Data

`r kable(phenotype_combined)`



`r if (nrow(proc_mappings) == 0) '# Manhattan plot'`
`r if (nrow(proc_mappings) == 0) 'The figure below shows the signifcance value for every single-nucleotide variant (SNV) present in the _C.elegans_ population, which are represented as black dots. The red line corresponds to the Bonferroni-corrected significance threshold, which is defined as $(-log_{10}\\frac{0.05}{\\#SNVs})$'`

`r if (nrow(proc_mappings) > 0) '# Mapping'`

```{r non-sig Manhattan Plot}
if(nrow(proc_mappings) == 0){
  
  readr::write_tsv(pr_mapping, "tables/non_sig_mapping.tsv")
  
  pr_mapping %>%
  ggplot2::ggplot(.) +
  ggplot2::aes(x = POS/1e6, y = log10p) +
  ggplot2::geom_point() +
  ggplot2::facet_grid(.~CHROM, scales = "free_x") +
  ggplot2::theme_bw() +
  ggplot2::geom_hline(aes(yintercept = bf), color = "#FF0000", size = 1)+
  theme_bw() +
  pub_theme + 
  theme(plot.margin = unit(c(0.0,0.5,0.5,0),"cm"),
        strip.background = element_blank(),
        axis.title.y = element_text(vjust=2.5),
        panel.border = element_rect(size=1, color = "black")) +
  ggplot2::labs(x = "Genomic Position (Mb)",
                y = expression(-log[10](p)))
}

```


```{r}
update_status("Processing Peaks")
if (nrow(proc_mappings) == 0) {
  chunk_run <- FALSE
  mult_chunk_run <- FALSE
} else {
  chunk_run <- TRUE
}

# Dummy peaks variable
peaks <- as.data.frame(c())
```


```{r, eval = chunk_run}
if (nrow(proc_mappings) != 0) {
  peaks <- na.omit(proc_mappings) %>%
  dplyr::distinct(peak_id, .keep_all = TRUE) %>%
  dplyr::select(marker, CHROM, startPOS, endPOS, log10p, trait) %>%
  dplyr::mutate(query = paste0(CHROM, ":",startPOS, "-",endPOS)) %>%
  dplyr::arrange(desc(log10p)) %>%
  dplyr::mutate(top3peaks = seq(1:n())) %>%
  dplyr::filter(top3peaks < 4) %>%
  dplyr::select(trait,peak_pos = marker, interval = query, peak_log10p = log10p)
} 
```

```{r}
if (nrow(proc_mappings) != 0){
  if(nrow(peaks) > 0){
    eval_cell_1 = TRUE
  } else {
    eval_cell_1 = FALSE
  } 
  
  if(nrow(peaks) > 1){
    eval_cell_2 = TRUE
  } else {
    eval_cell_2 = FALSE
  }
  
  if(nrow(peaks) > 2){
    eval_cell_3 = TRUE
  } else {
    eval_cell_3 = FALSE
  }
} else {
  eval_cell_1 = FALSE
  eval_cell_2 = FALSE
  eval_cell_3 = FALSE
}
```

`r if (eval_cell_1) '------------------------'`

`r if (eval_cell_1) '## Manhattan Plot'`

```{r Manplot, eval = chunk_run}
if (nrow(proc_mappings) != 0){
  mplot <- manplot(proc_mappings, "#666666")
  mplot[[1]] +
  theme_bw() +
  pub_theme + 
  theme(plot.margin = unit(c(0.0,0.5,0.5,0),"cm"),
        strip.background = element_blank(),
        #strip.text = element_blank(),
        axis.title.y = element_text(vjust=2.5),
        panel.border = element_rect(size=1, color = "black")) +
  ggplot2::labs(x = "Genomic Position (Mb)",
                y = expression(-log[10](p))) +
  theme(plot.title = ggplot2::element_blank())
}
```

`r if (eval_cell_1) 'A genome-wide representation of the association between variation in the _C.elegans_ population and your phenotype. The x-axis corresponds to genomic position with chromosome number indicated in the gray box above each facet. Dots represent single-nucleotide variant (SNV) present in the _C.elegans_ population. The y-axis corresponds to the level of significance for the association test. Blue dots represent SNVs that are above the defined significance threshold, which is the thick gray line. Red boxes surrounding blue dots represent the QTL confidence interval, which we define as plus/minus 50 SNVs from the last signifincant SNV. The default threshold for significance is the Bonferroni-corrected value $(-log_{10}\\frac{0.05}{\\#SNVs})$ and is usually around 5.5.'`

`r if (eval_cell_1) '------------------------'`

`r if (eval_cell_1) '## QTL Confidence Intervals'`

`r if(nrow(peaks) > 0 & chunk_run == T) { kable(peaks) } else { print("NO QTL IDENTIFIED") }`

`r if (eval_cell_1) '__QTL identified and corresponding confidence intervals__'`


`r if (eval_cell_1) '------------------------'`

`r if (eval_cell_1) '## QQ-plot'`

```{r QQplot, fig.height=12, fig.width=12, eval = chunk_run}
if (nrow(proc_mappings) != 0){
  log10ps <- distinct(proc_mappings, marker, .keep_all = TRUE)
  
  cegwas::qq_plot(log10ps$log10p) + scale_x_continuous(limits = c(0, NA)) +
    pub_theme
}
```

`r if (eval_cell_1) 'Quantile-quantile plot generated from the p-values from the association mapping. This plot shows the expected distribution of association test statistics on the x-axis compared to the observed values on the y-axis. Each dot corresponds to an individual SNV. SNVs that deviate from the _x=y_ line indicate that they are associated with your phenotype. Earlier separation from the identity line is indication of population stratification not being eliminated by the mapping algorithm.'`

`r if (eval_cell_1) '------------------------'`

`r if (eval_cell_1) '## Phenotype-by-Genotype'`

```{r PxGplot, eval = chunk_run}
if (nrow(proc_mappings) != 0){
  pg_plot <- pxg_plot(proc_mappings, color_strains = NA)
  pg_plot[[1]] +
    labs(y= args$trait_slug) +
    pub_theme + 
    theme(plot.margin = unit(c(0.0,0.5,0.5,0),"cm"),
        strip.background = element_blank(),
        #strip.text = element_blank(),
        axis.title.y = element_text(vjust=2.5),
        panel.border = element_rect(size=1, color = "black"))  +
    theme(legend.position = "none",
                      plot.title = ggplot2::element_blank())

}
```

`r if (eval_cell_1) 'Phenotypic distributions represeted as box plots are split by the genotype at the most significant SNV for a given QTL. The phenotype is on the y-axis and the genotype is on the x-axis. The chromosome and position for the plotted SNV are indicated in each the gray rectangles above the panels. REF refers to the N2 genotype and ALT refers to the variant phenotype. All of the SNVs in our data set are biallelic so these are the only two classes for any given SNV.'`

```{r, eval = chunk_run}

if (nrow(peaks) == 1) {
  mult_chunk_run <- FALSE
} else {
  mult_chunk_run <- TRUE
}

```

`r if (mult_chunk_run) '------------------------'`

`r if (mult_chunk_run) '## Linkage Disequilibrium'`

```{r LDplot, eval = mult_chunk_run, fig.align = "center", fig.width=14, fig.height=11}
if(nrow(peaks) > 1){
plot_peak_ld(proc_mappings) 
}
```

`r if (mult_chunk_run) 'Peak markers are the most statistically significant markers identified within a peak. The Linkage Disequilibrium (LD) measure being plotted is the correlation between peak markers $r = \\frac{-D}{\\sqrt{p(A) * p(a) * p(B) * p*(b)}}$, where D raw difference in frequency between the observed number of AB pairs and the expected number and A, B, a, and b refer to alleles at the loci. An LD value of 0.8 or higher suggests that the two peak markers are not segregating randomly.'`

```{r interval_variants, eval = chunk_run}
if(nrow(proc_mappings) != 0){
  interval_variants <- process_correlations(variant_correlation(proc_mappings, quantile_cutoff_high = 0.75, quantile_cutoff_low = 0.25, condition_trait = F))
  
  readr::write_tsv(interval_variants, "tables/interval_variants.tsv")
  
  # Condense Interval Variants File
  interval_variants %>% 
    dplyr::select(CHROM, POS, gene_id, num_alt_allele, num_strains, corrected_spearman_cor) %>%
    dplyr::distinct(.keep_all = T) %>%
    readr::write_tsv("tables/interval_variants_db.tsv")
}
```

```{r}
update_status("Transferring Data")
```
